---
title: "Prior"
author: "Raymond"
format: html
editor: visual
---

## EDA

I use data from the tail-base, mid-center and nose to get a prior distribution of the three pairs of body part distances. From the histograms, both the actual data and manually labelled distances of body parts seem roughly Gaussian if you discard obvious outliers (mistakes). It seems reasonable to model the distances as Gaussian distributions with the MLE plug-in estimators of the sample mean and standard deviation using manually labelled data.

I also considered using the ratios of distances, but this seems less stable so it was not used in the final model.

```{r, message = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(purrr)
library(plotly)
library(tidyverse)
library(tidyr)
coords = read_csv("predictions_aligned_40.csv")
str(coords)
```

```{r}
library(dplyr)
coords_wide <- coords |>
  filter(y < 730 | is.na(y)) |>
  select(frame, instance, node, x, y) |>
  pivot_wider(
    names_from = node,
    values_from = c(x, y),
    names_sep = "_"
  )
coords_mid = coords_wide |>
  mutate(dtn = sqrt((x_Nose - `x_Tail-base`)^2 + 
                      (y_Nose - `y_Tail-base`)^2)) |>
  mutate(dtm = sqrt((`x_Mid-center` - `x_Tail-base`)^2 + 
                    (`y_Mid-center` - `y_Tail-base`)^2)) |>
  mutate(dnm = sqrt((`x_Mid-center` - x_Nose)^2 + 
                    (`y_Mid-center` - y_Nose)^2)) |>
  mutate(dist_ratio = dtm / dnm) 

```

Examine distributions of real data tail-nose distances:

```{r}
ggplot() +
  geom_histogram(data = coords_mid, aes(x = log(dist_ratio), y = after_stat(density))) +
  labs(title = "Real Distance Ratios")
ggplot() +
  geom_histogram(data = coords_mid, aes(x = dtn, y = after_stat(density))) +
  labs(title = "Real Tail Nose Distances")
```

```{r, message  = FALSE}
manual = read_csv("predictions_manual.csv")
manual = manual |>
  filter(score == 1) |>
  filter(frame != 308) |>
  filter(frame != 310) |>
  filter(frame != 3511) |>
  filter(frame != 581)
manual$instance =  rep(rep(0:1, each = 8), 40)
```

Examine distributions of manually labelled body part distances:

```{r}
manual_wide <- manual |>
  select(frame, instance, node, x, y) |>
  pivot_wider(
    names_from = node,
    values_from = c(x, y),
    names_sep = "_"
  )
manual_mid = manual_wide |>
  mutate(dtn = sqrt((x_Nose - `x_Tail-base`)^2 + 
                      (y_Nose - `y_Tail-base`)^2)) |>
  mutate(dtm = sqrt((`x_Mid-center` - `x_Tail-base`)^2 + 
                    (`y_Mid-center` - `y_Tail-base`)^2)) |>
  mutate(dnm = sqrt((`x_Mid-center` - x_Nose)^2 + 
                    (`y_Mid-center` - y_Nose)^2)) |>
  mutate(dist_ratio = dtm / dnm)
ggplot() +
  geom_histogram(data = manual_mid, aes(x = dtm, y = after_stat(density))) +
  labs(title = "Tail Mid-center Distances")
ggplot() +
  geom_histogram(data = manual_mid, aes(x = dtn, y = after_stat(density))) +
  labs(title = "Tail Nose Distances")
ggplot() +
  geom_histogram(data = manual_mid, aes(x = dnm, y = after_stat(density))) +
  labs(title = "Nose Mid-center Distances")
```

```{r}
library(dplyr)
library(tidyr)

library(dplyr)

library(dplyr)

manual_angles <- manual_wide %>%
  # vectors from Mid -> Nose and Mid -> Tail-base
  mutate(
    vx_mn = x_Nose - `x_Mid-center`,
    vy_mn = y_Nose - `y_Mid-center`,
    vx_mt = `x_Tail-base` - `x_Mid-center`,
    vy_mt = `y_Tail-base` - `y_Mid-center`,
    # angles
    ang_mn = atan2(vy_mn, vx_mn),
    ang_mt = atan2(vy_mt, vx_mt),
    angle_mid = (ang_mn - ang_mt + pi) %% (2*pi) - pi
  ) %>%
  summarise(
    # circular mean relative to pi
    mean_cos = mean(cos(angle_mid - pi), na.rm = TRUE),
    mean_sin = mean(sin(angle_mid - pi), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    circ_mean_rad = atan2(mean_sin, mean_cos) + pi,  # mean angle
    R = sqrt(mean_cos^2 + mean_sin^2),               # mean resultant length
    circ_sd_rad = sqrt(pmax(0, -2 * log(R))),        # circular SD
    circ_sd_deg = circ_sd_rad * 180 / pi
  ) %>%
  select(circ_mean_rad, circ_sd_rad, circ_sd_deg)

circ_sd = manual_angles$circ_sd_rad
circ_mean = manual_angles$circ_mean_rad

```

```{r}
pri_dtm_mean = mean(manual_mid$dtm)
pri_dtm_sd = sd(manual_mid$dtm)
pri_dtn_mean = mean(manual_mid$dtn)
pri_dtn_sd = sd(manual_mid$dtn)
pri_dnm_mean = mean(manual_mid$dnm)
pri_dnm_sd = sd(manual_mid$dnm)

pri_list <- list(
  circ_mean = manual_angles$circ_mean_rad,
  circ_sd   = manual_angles$circ_sd_rad,
  
  pri_dtm_mean = mean(manual_mid$dtm, na.rm = TRUE),
  pri_dtm_sd   = sd(manual_mid$dtm, na.rm = TRUE),
  
  pri_dtn_mean = mean(manual_mid$dtn, na.rm = TRUE),
  pri_dtn_sd   = sd(manual_mid$dtn, na.rm = TRUE),
  
  pri_dnm_mean = mean(manual_mid$dnm, na.rm = TRUE),
  pri_dnm_sd   = sd(manual_mid$dnm, na.rm = TRUE)
)

# --- Save as RDS file ---
saveRDS(pri_list, file = "priors_list.rds")

```

Examine the data in a loop in frame order, but with real time updates of the dataframe before going to the next frame. For each frame, try every combination of the coordinates for instance 0 (So for tail-base, mid-center, and nose nodes, there will be 8 combinations, including potential NAs). Instance 1 will get the remaining coordinates. Then, for each combination compute a score based on average likelihood of the 3 distances of nodes according to prior (removing NAs) as well as weighted average distance to the previous frame (the weight decreases as the time gap largens). Update the two instances with the combinations accepting lowest score below threshold, if no score is below threshold, then keep the coordinates of the previous frame. After a full pass through the data, mark the frames where a score was accepted as reliable. The decaying weight allows the instance where the track has gone cold for many frames to jump to the other instance once it reappears

```{r}
preprocess_coords <- function(coords_sort) {

  coords_sort %>%
    group_by(frame) %>%
    group_modify(~ {
      df <- .
      if (nrow(df) <= 2) return(df)  # nothing to do

      # Compute pairwise distances between mid-centers
      mids <- df %>% select(x_Mid_center = `x_Mid-center`, y_Mid_center = `y_Mid-center`)
      dists <- as.matrix(dist(mids))  # 3x3 symmetric

      # Get unique pairs
      pairs <- expand.grid(i = 1:3, j = 1:3)
      pairs <- pairs[pairs$i < pairs$j, ]
      pairs$dist <- mapply(function(i,j) dists[i,j], pairs$i, pairs$j)
      closest_pair <- pairs %>% slice_min(dist, with_ties = FALSE)

      # Remove the second of the closest pair
      remove_idx <- closest_pair$j
      df_clean <- df[-remove_idx, ]

      # Relabel instance as 0 and 1 using nrow()
      df_clean <- df_clean %>%
        mutate(instance = 0:(nrow(df_clean)-1))

      df_clean
    }) %>%
    ungroup()
}


coords_sort = preprocess_coords(coords_mid |> arrange(frame))
saveRDS(coords_sort, "SLEAP_pred.rds")
```

Helper Functions for tracking:

```{r}
process_forward_combos <- function(coords_sort, r, priors, threshold, motion_alpha = 1, big_penalty = 1e6, too_close_thresh = 10, w_ang = 1,
decay_rate = 0.8, decay_ang = 0.5, verbose = FALSE) {
  # --- helpers ---
  nll <- function(d, mean, sd) {
    if (length(d) == 0 || is.na(d)) return(0)
    0.5 * ((d - mean) / sd)^2 + log(sd)
  }
  dist_xy <- function(a, b) {
    if (length(a) != 2 || length(b) != 2) return(NA_real_)
    if (any(is.na(a)) || any(is.na(b))) return(NA_real_)
    sqrt((a[1] - b[1])^2 + (a[2] - b[2])^2)
  }
  # shortest signed angle difference (returns value in (-pi, pi])
  ang_diff <- function(a, b) {
    d <- a - b
    atan2(sin(d), cos(d))
  }

  # angle at `mid` between vectors (p - mid) and (r - mid)
  angle_at_mid <- function(p, mid, r) {
    v1 <- p - mid
    v2 <- r - mid
    if (any(is.na(v1)) || any(is.na(v2))) return(NA_real_)
    a1 <- atan2(v1[2], v1[1])
    a2 <- atan2(v2[2], v2[1])
    abs(ang_diff(a1, a2)) # [0, pi]
  }

  # Normal NLL for angle (wrap-safe)
  angle_nll <- function(angle_rad, prior_mean = pi, prior_sd = 1/pi) {
    if (is.na(angle_rad)) return(NA_real_)
    d <- ang_diff(angle_rad, prior_mean)
    0.5 * (d / prior_sd)^2 + log(prior_sd)
  }

  # --- checks ---
  required_cols <- c("frame", "instance", "x_Nose", "y_Nose",
                     "x_Mid-center", "y_Mid-center",
                     "x_Tail-base", "y_Tail-base")
  stopifnot(all(required_cols %in% colnames(coords_sort)))
  frames_all <- sort(unique(coords_sort$frame))
  if (! r %in% frames_all)
    stop("start frame r not present in coords_sort")

  combos <- expand.grid(nose_choice = 1:2,
                        mid_choice  = 1:2,
                        tail_choice = 1:2,
                        stringsAsFactors = FALSE)
  frames_seq <- frames_all[frames_all >= r]

  # --- initialize from frame r ---
  fr_r_rows <- coords_sort %>% dplyr::filter(frame == r) %>% dplyr::arrange(instance)
  if (nrow(fr_r_rows) == 0) stop("No rows at start frame r")
  if (nrow(fr_r_rows) == 1 && f != r) {
    pad_row <- fr_r_rows[1, , drop = FALSE]; pad_row[,] <- NA
    fr_r_rows <- dplyr::bind_rows(fr_r_rows, pad_row)
  } else {
    fr_r_rows <- fr_r_rows[1:2, , drop = FALSE]
  }

  # prev accepted info — per-part previous frames (consistent names)
  prev_frame_nose0 <- prev_frame_mid0 <- prev_frame_tail0 <- r
  prev_frame_nose1 <- prev_frame_mid1 <- prev_frame_tail1 <- r

  prev_nose0 <- c(fr_r_rows[1, "x_Nose"], fr_r_rows[1, "y_Nose"]) %>% as.numeric()
  prev_mid0  <- c(fr_r_rows[1, "x_Mid-center"], fr_r_rows[1, "y_Mid-center"]) %>% as.numeric()
  prev_tail0 <- c(fr_r_rows[1, "x_Tail-base"], fr_r_rows[1, "y_Tail-base"]) %>% as.numeric()
  prev_nose1 <- c(fr_r_rows[2, "x_Nose"], fr_r_rows[2, "y_Nose"]) %>% as.numeric()
  prev_mid1  <- c(fr_r_rows[2, "x_Mid-center"], fr_r_rows[2, "y_Mid-center"]) %>% as.numeric()
  prev_tail1 <- c(fr_r_rows[2, "x_Tail-base"], fr_r_rows[2, "y_Tail-base"]) %>% as.numeric()
  pred_mid0 = prev_mid0
  pred_mid1 = prev_mid1
  pred_nose0 = prev_nose0
  pred_nose1 = prev_nose1
  pred_tail0 = prev_tail0
  pred_tail1 = prev_tail1

  out <- vector("list", length(frames_seq))
  row_index <- 0

  # --- loop over frames ---
  for (f in frames_seq) {
    row_index <- row_index + 1
    fr_rows <- coords_sort %>% dplyr::filter(frame == f) %>% dplyr::arrange(instance)


    # no detections: carry forward
    if (nrow(fr_rows) == 0) {
      out[[row_index]] <- tibble::tibble(
        frame = f, best_combo = NA_integer_,
        score0 = NA_real_, score1 = NA_real_,
        inst0_mid_x = prev_mid0[1], inst0_mid_y = prev_mid0[2],
        inst1_mid_x = prev_mid1[1], inst1_mid_y = prev_mid1[2],
        inst0_nose_x = prev_nose0[1], inst0_nose_y = prev_nose0[2],
        inst1_nose_x = prev_nose1[1], inst1_nose_y = prev_nose1[2],
        inst0_tail_x = prev_tail0[1], inst0_tail_y = prev_tail0[2],
        inst1_tail_x = prev_tail1[1], inst1_tail_y = prev_tail1[2],
        reliability_mid0 = FALSE, reliability_mid1 = FALSE,
        reliability_nose0 = FALSE, reliability_nose1 = FALSE,
        reliability_tail0 = FALSE, reliability_tail1 = FALSE
      )
      next
    }

    if (nrow(fr_rows) == 1) {
      pad_row <- fr_rows[1, , drop = FALSE]; pad_row[,] <- NA
      fr_rows <- dplyr::bind_rows(fr_rows, pad_row)
    } else {
      fr_rows <- fr_rows[1:2, , drop = FALSE]
    }

    # helpers: too-close and fix
    too_close <- function(p1, p2, thresh) {
      if (any(is.na(p1)) || any(is.na(p2))) return(FALSE)
      sqrt(sum((p1 - p2)^2)) < thresh
    }
    fix_too_close <- function(cands, too_close_thresh) {
      if (length(cands) >= 2) {
        if (too_close(cands[[1]], cands[[2]], too_close_thresh)) {
          mean_pt <- colMeans(rbind(cands[[1]], cands[[2]]), na.rm = TRUE)
          cands[[1]] <- mean_pt
          cands[[2]] <- c(NA, NA)
        }
      }
      cands
    }

    # candidate lists (with cleaning)
    nose_cands <- list(
      c(fr_rows[["x_Nose"]][1], fr_rows[["y_Nose"]][1]) %>% as.numeric(),
      c(fr_rows[["x_Nose"]][2], fr_rows[["y_Nose"]][2]) %>% as.numeric()
    ) %>% fix_too_close(too_close_thresh)

    mid_cands <- list(
      c(fr_rows[["x_Mid-center"]][1], fr_rows[["y_Mid-center"]][1]) %>% as.numeric(),
      c(fr_rows[["x_Mid-center"]][2], fr_rows[["y_Mid-center"]][2]) %>% as.numeric()
    ) %>% fix_too_close(too_close_thresh)

    tail_cands <- list(
      c(fr_rows[["x_Tail-base"]][1], fr_rows[["y_Tail-base"]][1]) %>% as.numeric(),
      c(fr_rows[["x_Tail-base"]][2], fr_rows[["y_Tail-base"]][2]) %>% as.numeric()
    ) %>% fix_too_close(too_close_thresh)

    # prepare combo scores table
    combo_scores <- tibble::tibble(
      combo = seq_len(nrow(combos)),
      score0 = NA_real_, score1 = NA_real_,
      mid0x = NA_real_, mid0y = NA_real_,
      mid1x = NA_real_, mid1y = NA_real_,
      nose0x = NA_real_, nose0y = NA_real_,
      nose1x = NA_real_, nose1y = NA_real_,
      tail0x = NA_real_, tail0y = NA_real_,
      tail1x = NA_real_, tail1y = NA_real_
    )

    fill_with_prev <- function(curr, prev) {
      if (any(is.na(curr))) return(prev)
      curr
    }

    # --- compute scores for every combo ---
    for (ci in seq_len(nrow(combos))) {
      nc <- combos$nose_choice[ci]
      mc <- combos$mid_choice[ci]
      tc <- combos$tail_choice[ci]

      # raw candidates for instance 0 (choices) and instance 1 (the other ones)
      nose0 <- nose_cands[[nc]];   nose1 <- nose_cands[[3 - nc]]
      mid0  <- mid_cands[[mc]];    mid1  <- mid_cands[[3 - mc]]
      tail0 <- tail_cands[[tc]];   tail1 <- tail_cands[[3 - tc]]

      # replace NA with previous frame coords
      nose0 <- fill_with_prev(nose0, pred_nose0);   mid0  <- fill_with_prev(mid0,  pred_mid0);   tail0 <- fill_with_prev(tail0, pred_tail0)
      nose1 <- fill_with_prev(nose1, pred_nose1);   mid1  <- fill_with_prev(mid1,  pred_mid1);   tail1 <- fill_with_prev(tail1, pred_tail1)

      # distances for instance 0 and 1
      d_nm0 <- dist_xy(nose0, mid0); d_tm0 <- dist_xy(tail0, mid0); d_tn0 <- dist_xy(tail0, nose0)
      d_nm1 <- dist_xy(nose1, mid1); d_tm1 <- dist_xy(tail1, mid1); d_tn1 <- dist_xy(tail1, nose1)

      # frame gaps per part (use the per-part prev_frame_* variables)
      frame_gap_nose0 <- max(1, f - prev_frame_nose0)
      frame_gap_mid0  <- max(1, f - prev_frame_mid0)
      frame_gap_tail0 <- max(1, f - prev_frame_tail0)

      frame_gap_nose1 <- max(1, f - prev_frame_nose1)
      frame_gap_mid1  <- max(1, f - prev_frame_mid1)
      frame_gap_tail1 <- max(1, f - prev_frame_tail1)

      # per-part NLLs with decay
      nll_nm0 <- if (!is.na(d_nm0)) nll(d_nm0, priors$NM["mean"], priors$NM["sd"]) * decay_rate^(max(frame_gap_nose0, frame_gap_mid0) - 1) else NA_real_
      nll_tm0 <- if (!is.na(d_tm0)) nll(d_tm0, priors$TM["mean"], priors$TM["sd"]) * decay_rate^(max(frame_gap_tail0, frame_gap_mid0) - 1) else NA_real_
      nll_tn0 <- if (!is.na(d_tn0)) nll(d_tn0, priors$TN["mean"], priors$TN["sd"]) * decay_rate^(max(frame_gap_tail0, frame_gap_nose0) - 1) else NA_real_

      nll_nm1 <- if (!is.na(d_nm1)) nll(d_nm1, priors$NM["mean"], priors$NM["sd"]) * decay_rate^(max(frame_gap_nose1, frame_gap_mid1) - 1) else NA_real_
      nll_tm1 <- if (!is.na(d_tm1)) nll(d_tm1, priors$TM["mean"], priors$TM["sd"]) * decay_rate^(max(frame_gap_tail1, frame_gap_mid1) - 1) else NA_real_
      nll_tn1 <- if (!is.na(d_tn1)) nll(d_tn1, priors$TN["mean"], priors$TN["sd"]) * decay_rate^(max(frame_gap_tail1, frame_gap_nose1) - 1) else NA_real_

      avg_nll0 <- if (all(is.na(c(nll_nm0, nll_tm0, nll_tn0)))) big_penalty else mean(c(nll_nm0, nll_tm0, nll_tn0), na.rm = TRUE)
      avg_nll1 <- if (all(is.na(c(nll_nm1, nll_tm1, nll_tn1)))) big_penalty else mean(c(nll_nm1, nll_tm1, nll_tn1), na.rm = TRUE)

      # motion penalties (per part)
      dists0 <- c(nose = dist_xy(prev_nose0, nose0), mid = dist_xy(prev_mid0, mid0), tail = dist_xy(prev_tail0, tail0))
      dists1 <- c(nose = dist_xy(prev_nose1, nose1), mid = dist_xy(prev_mid1, mid1), tail = dist_xy(prev_tail1, tail1))

      gaps0 <- c(nose = frame_gap_nose0, mid = frame_gap_mid0, tail = frame_gap_tail0)
      gaps1 <- c(nose = frame_gap_nose1, mid = frame_gap_mid1, tail = frame_gap_tail1)

      penalties0 <- mapply(function(mot, gap) if (!is.na(mot) && mot > (gap  * motion_alpha)) (mot - gap * motion_alpha) 
                           else if(mot > max((gap - 1), 2) * motion_alpha) threshold / 2                     
                           else 0, dists0, gaps0)
      penalties1 <- mapply(function(mot, gap) if (!is.na(mot) && mot > (gap * motion_alpha)) (mot - gap * motion_alpha) 
                           else if(mot > max((gap - 1), 2) * motion_alpha) threshold / 2
                           else 0, dists1, gaps1)

      motion_pen0 <- sum(penalties0)
      motion_pen1 <- sum(penalties1)

      # angle priors
      ang0 <- angle_at_mid(nose0, mid0, tail0)
      ang1 <- angle_at_mid(nose1, mid1, tail1)
      nll_ang0 <- angle_nll(ang0, prior_mean = circ_mean, prior_sd = circ_sd) * decay_ang ^ (max(frame_gap_tail0, frame_gap_nose0, frame_gap_mid0) - 1)
      nll_ang1 <- angle_nll(ang1, prior_mean = circ_mean, prior_sd = circ_sd) * decay_ang ^ (max(frame_gap_tail1, frame_gap_nose1, frame_gap_mid1) - 1)
      
      # vectors between mid0 and mid1
prev_mid_vec <- prev_mid0 - prev_mid1
curr_mid_vec <- mid0 - mid1

angle_change_mid <- NA_real_

# compute angle between previous and current mid vector
if (!any(is.na(c(prev_mid_vec, curr_mid_vec)))) {
  cosang <- sum(prev_mid_vec * curr_mid_vec) / 
            (sqrt(sum(prev_mid_vec^2)) * sqrt(sum(curr_mid_vec^2)))
  cosang <- max(min(cosang, 1), -1)  # clamp to [-1, 1] to avoid numerical issues
  angle_change_mid <- acos(cosang)
}

# penalize if angle change > 90°
angle_pen_mid <- if (!is.na(angle_change_mid) && angle_change_mid > pi/2) threshold / 2 else 0


# --- combine into total scores ---
combo_scores$score0[ci] <- avg_nll0 + motion_pen0 + w_ang * nll_ang0 + angle_pen_mid
combo_scores$score1[ci] <- avg_nll1 + motion_pen1 + w_ang * nll_ang1 + angle_pen_mid


      # store coordinates
      combo_scores$mid0x[ci]   <- mid0[1];  combo_scores$mid0y[ci]   <- mid0[2]
      combo_scores$mid1x[ci]   <- mid1[1];  combo_scores$mid1y[ci]   <- mid1[2]
      combo_scores$nose0x[ci]  <- nose0[1]; combo_scores$nose0y[ci]  <- nose0[2]
      combo_scores$nose1x[ci]  <- nose1[1]; combo_scores$nose1y[ci]  <- nose1[2]
      combo_scores$tail0x[ci]  <- tail0[1]; combo_scores$tail0y[ci]  <- tail0[2]
      combo_scores$tail1x[ci]  <- tail1[1]; combo_scores$tail1y[ci]  <- tail1[2]
    } # end combos loop

    # mid distance (for filtering too-close midpoints)
    combo_scores <- combo_scores |> dplyr::mutate(mid_dist = sqrt((mid0x - mid1x)^2 + (mid0y - mid1y)^2)) |> dplyr::mutate(mid_sum0 = sqrt((mid0x - prev_mid0[1])^2  + (mid0y - prev_mid0[2])^2)) |>
      dplyr::mutate(mid_sum1 = sqrt((mid1x - prev_mid1[1])^2  + (mid1y - prev_mid1[2])^2))
    tol <- 1e-6

    dist0_prev1 <- sqrt((combo_scores$mid0x - pred_mid1[1])^2 + (combo_scores$mid0y - pred_mid1[2])^2)
    dist1_prev0 <- sqrt((combo_scores$mid1x - pred_mid0[1])^2 + (combo_scores$mid1y - pred_mid0[2])^2)
    combo_scores <- combo_scores |>
  dplyr::mutate(
    score0 = score0 + ifelse(dist0_prev1 < mid_sum0 - tol, threshold / 4 , 0),
    score1 = score1 + ifelse(dist1_prev0 < mid_sum1 - tol, threshold / 4, 0)
  )

    valid_both <- combo_scores %>%
      dplyr::filter(!is.na(score0), !is.na(score1), score0 < threshold, score1 < threshold, mid_dist >= motion_alpha, mid_sum0 / 3 <=  dist0_prev1, mid_sum1 / 3 <= dist1_prev0
)
    
    valid0 <- combo_scores %>% dplyr::filter(!is.na(score0), mid_dist > motion_alpha, score0 < threshold, mid_sum0 / 3 <=  dist0_prev1)
    
    valid1 <- combo_scores %>% dplyr::filter(!is.na(score1), mid_dist > motion_alpha, score1 < threshold, mid_sum1 / 3 <= dist1_prev0)
    # decision logic
  
chosen <- NULL; accept0 <- FALSE; accept1 <- FALSE
    tol <- 1e-6

    if (nrow(valid_both) > 0) {
      comb_dist <- sqrt((valid_both$mid0x - prev_mid0[1])^2 + (valid_both$mid0y - prev_mid0[2])^2) +
                   sqrt((valid_both$mid1x - prev_mid1[1])^2 + (valid_both$mid1y - prev_mid1[2])^2)
      chosen <- valid_both[which.min(comb_dist), , drop = FALSE]
      accept0 <- TRUE; accept1 <- TRUE
    } else {
      # compute distances for single-instance choices
      if (nrow(valid0) > 0) {
        d0 <- sqrt((valid0$mid0x - prev_mid0[1])^2 + (valid0$mid0y - prev_mid0[2])^2)
      } else d0 <- numeric(0)
      if (nrow(valid1) > 0) {
        d1 <- sqrt((valid1$mid1x - prev_mid1[1])^2 + (valid1$mid1y - prev_mid1[2])^2)
      } else d1 <- numeric(0)

      valid0_pos <- if (length(d0) > 0) valid0[d0 > tol, , drop = FALSE] else valid0[0, , drop = FALSE]
      valid1_pos <- if (length(d1) > 0) valid1[d1 > tol, , drop = FALSE] else valid1[0, , drop = FALSE]
      d0_pos <- if (length(d0) > 0) d0[d0 > tol] else numeric(0)
      d1_pos <- if (length(d1) > 0) d1[d1 > tol] else numeric(0)

      if (nrow(valid0_pos) > 0 && nrow(valid1_pos) > 0) {
        best0 <- valid0_pos[which.min(d0_pos), , drop = FALSE]
        best1 <- valid1_pos[which.min(d1_pos), , drop = FALSE]
        if (min(d0_pos) <= min(d1_pos)) {
          chosen <- best0; accept0 <- TRUE
        } else {
          chosen <- best1; accept1 <- TRUE
        }
      } else if (nrow(valid0_pos) > 0) {
        chosen <- valid0_pos[which.min(d0_pos), , drop = FALSE]; accept0 <- TRUE
      } else if (nrow(valid1_pos) > 0) {
        chosen <- valid1_pos[which.min(d1_pos), , drop = FALSE]; accept1 <- TRUE
      } else {
        # fallback: carry-forward previous positions
        chosen <- tibble::tibble(
          combo = NA_integer_,
          mid0x = prev_mid0[1], mid0y = prev_mid0[2],
          mid1x = prev_mid1[1], mid1y = prev_mid1[2],
          nose0x = prev_nose0[1], nose0y = prev_nose0[2],
          nose1x = prev_nose1[1], nose1y = prev_nose1[2],
          tail0x = prev_tail0[1], tail0y = prev_tail0[2],
          tail1x = prev_tail1[1], tail1y = prev_tail1[2],
          score0 = Inf, score1 = Inf
        )
      }
    }

          # initialize per-part reliability flags
    reliability_mid0 <- reliability_mid1 <- FALSE
    reliability_nose0 <- reliability_nose1 <- FALSE
    reliability_tail0 <- reliability_tail1 <- FALSE

# ---------- instance 0 ----------
if (accept0) {
  # mid0
  if (!any(is.na(chosen[c("mid0x", "mid0y")]))) {
    moved <- !isTRUE(all.equal(prev_mid0, c(chosen$mid0x, chosen$mid0y), tolerance = tol))
    reliability_mid0 <- as.logical(moved)
    prev_frame_mid0 <- f
    if (reliability_mid0) {
      prev_mid0 <- c(chosen$mid0x, chosen$mid0y)
      pred_mid0 = prev_mid0
    }
  }

  # nose0
  if (!any(is.na(chosen[c("nose0x", "nose0y")]))) {
    moved <- !isTRUE(all.equal(prev_nose0, c(chosen$nose0x, chosen$nose0y), tolerance = tol))
    reliability_nose0 <- as.logical(moved)
    prev_frame_nose0 <- f
    if (reliability_nose0) {
      prev_nose0 <- c(chosen$nose0x, chosen$nose0y)
      pred_nose0 = prev_nose0
    }
  }

  # tail0
  if (!any(is.na(chosen[c("tail0x", "tail0y")]))) {
    moved <- !isTRUE(all.equal(prev_tail0, c(chosen$tail0x, chosen$tail0y), tolerance = tol))
    reliability_tail0 <- as.logical(moved)
     prev_frame_tail0 <- f
    if (reliability_tail0) {
      prev_tail0 <- c(chosen$tail0x, chosen$tail0y)
      pred_tail0 = prev_tail0
    }
  }
}

# ---------- instance 1 ----------
if (accept1) {
  # mid1
  if (!any(is.na(chosen[c("mid1x", "mid1y")]))) {
    moved <- !isTRUE(all.equal(prev_mid1, c(chosen$mid1x, chosen$mid1y), tolerance = tol))
    reliability_mid1 <- as.logical(moved)
    prev_frame_mid1 <- f
    if (reliability_mid1) {
      prev_mid1 <- c(chosen$mid1x, chosen$mid1y)
      pred_mid1 = prev_mid1
    }
  }

  # nose1
  if (!any(is.na(chosen[c("nose1x", "nose1y")]))) {
    moved <- !isTRUE(all.equal(prev_nose1, c(chosen$nose1x, chosen$nose1y), tolerance = tol))
    reliability_nose1 <- as.logical(moved)
    prev_frame_nose1 <- f
    if (reliability_nose1) {
      prev_nose1 <- c(chosen$nose1x, chosen$nose1y)
      pred_nose1 = prev_nose1
    }
  }

  # tail1
  if (!any(is.na(chosen[c("tail1x", "tail1y")]))) {
    moved <- !isTRUE(all.equal(prev_tail1, c(chosen$tail1x, chosen$tail1y), tolerance = tol))
    reliability_tail1 <- as.logical(moved)
    prev_frame_tail1 <- f
    if (reliability_tail1) {
      prev_tail1 <- c(chosen$tail1x, chosen$tail1y)
      pred_tail1 = prev_tail1
    }
  }
}
    # ---------- frame 0 (start frame r) ----------
if (f == r) {
  # Only mark as reliable if the coordinates are non-NA
  reliability_mid0  <- all(!is.na(prev_mid0))
  reliability_mid1  <- all(!is.na(prev_mid1))
  reliability_nose0 <- all(!is.na(prev_nose0))
  reliability_nose1 <- all(!is.na(prev_nose1))
  reliability_tail0 <- all(!is.na(prev_tail0))
  reliability_tail1 <- all(!is.na(prev_tail1))
}
update_pred <- function(prev_pt, unrel_pt, cands, reliable, motion_thresh0, motion_thresh1, other_prev = NULL) {
  if (reliable) {
    # reliable → just use previous
    return(prev_pt)
  }

  # --- filter out candidates that are too close to the other instance ---
  if (!is.null(other_prev) && !any(is.na(other_prev))) {
    cands <- Filter(function(c) {
      if (any(is.na(c))) return(FALSE)
      sqrt(sum((c - other_prev)^2)) >= motion_thresh1
    }, cands)
  }

  if (length(cands) == 0) {
    return(unrel_pt)
  }

  # --- compute distances from unrel_pt ---
  dists <- sapply(cands, function(c) {
    if (any(is.na(c)) || any(is.na(unrel_pt))) return(NA_real_)
    sqrt(sum((c - unrel_pt)^2))
  })

  # remove NA and zero distances
  valid_idx <- which(!is.na(dists) & dists > 0)
  if (length(valid_idx) == 0) {
    return(unrel_pt)
  }

  dists <- dists[valid_idx]
  cands <- cands[valid_idx]

  # if all candidates are too far → keep unrel_pt
  if (min(dists) >= motion_thresh0) {
    return(unrel_pt)
  }

  # pick closest valid candidate
  cands[[which.min(dists)]]
}

pred_mid0  <- update_pred(prev_mid0,  pred_mid0,  mid_cands,  reliability_mid0,
                          motion_alpha * 3, motion_alpha * 2, other_prev = pred_mid1)
pred_mid1  <- update_pred(prev_mid1,  pred_mid1,  mid_cands,  reliability_mid1,
                          motion_alpha * 3, motion_alpha * 2, other_prev = pred_mid0)

pred_nose0 <- update_pred(prev_nose0, pred_nose0, nose_cands, reliability_nose0,
                          motion_alpha * 3,  motion_alpha * 2, other_prev = pred_nose1)
pred_nose1 <- update_pred(prev_nose1, pred_nose1, nose_cands, reliability_nose1,
                          motion_alpha * 3, motion_alpha * 2, other_prev = pred_nose0)

pred_tail0 <- update_pred(prev_tail0, pred_tail0, tail_cands, reliability_tail0,
                          motion_alpha * 3, motion_alpha * 2, other_prev = pred_tail1)
pred_tail1 <- update_pred(prev_tail1, pred_tail1, tail_cands, reliability_tail1,
                          motion_alpha * 3, motion_alpha * 2, other_prev = pred_tail0)



# --- save result ---
out[[row_index]] <- tibble::tibble(
  frame = f,
  best_combo = if (!is.null(chosen$combo)) chosen$combo else NA_integer_,
  score0 = if (!is.null(chosen$score0)) chosen$score0 else NA_real_,
  score1 = if (!is.null(chosen$score1)) chosen$score1 else NA_real_,
  inst0_mid_x  = pred_mid0[1],  inst0_mid_y  = pred_mid0[2],
  inst1_mid_x  = pred_mid1[1],  inst1_mid_y  = pred_mid1[2],
  inst0_nose_x = pred_nose0[1], inst0_nose_y = pred_nose0[2],
  inst1_nose_x = pred_nose1[1], inst1_nose_y = pred_nose1[2],
  inst0_tail_x = pred_tail0[1], inst0_tail_y = pred_tail0[2],
  inst1_tail_x = pred_tail1[1], inst1_tail_y = pred_tail1[2],
  reliability_mid0  = reliability_mid0,  reliability_mid1  = reliability_mid1,
  reliability_nose0 = reliability_nose0, reliability_nose1 = reliability_nose1,
  reliability_tail0 = reliability_tail0, reliability_tail1 = reliability_tail1
)


    if (verbose && (f %% 500 == 0)) {
      message("Frame ", f, ": score0=", signif(ifelse(is.null(chosen$score0), NA, chosen$score0), 4),
              ", score1=", signif(ifelse(is.null(chosen$score1), NA, chosen$score1), 4),
              ", accept0=", accept0, ", accept1=", accept1)
    }
  } # frames loop

  dplyr::bind_rows(out)
}

```

Usage of the helper functions to track instances:

```{r}
priors <- list(
  NM = c(mean = pri_dnm_mean, sd = pri_dnm_sd),
  TM = c(mean = pri_dtm_mean, sd = pri_dtm_sd),
  TN = c(mean = pri_dtn_mean, sd = pri_dtn_sd)
)

res <- process_forward_combos(
  coords_sort = coords_sort,
  r = 0,
  priors = priors,
  threshold = 35,
  too_close_thresh = 40,
  motion_alpha = 50,
  w_ang = 9,
  big_penalty = 1e5,
  decay_rate = 0.8,
  decay_ang = 0.9,
  verbose = TRUE
)


```

```{r}
library(dplyr)
library(ggplot2)

# Helper function to reshape and plot with part-specific reliability
plot_part_axis <- function(res, part, axis = c("x","y"), frame_range = c(5000, 6000)) {
  axis <- match.arg(axis)
  
  # column names
  col0 <- paste0("inst0_", part, "_", axis)
  col1 <- paste0("inst1_", part, "_", axis)
  
  # part-specific reliabilities
  rel0 <- paste0("reliability_", part, "0")
  rel1 <- paste0("reliability_", part, "1")
  
  # reshape for plotting
  df0 <- res %>%
    filter(frame >= frame_range[1], frame <= frame_range[2]) %>%
    transmute(frame, value = .data[[col0]], reliable = .data[[rel0]], instance = "inst0")
  
  df1 <- res %>%
    filter(frame >= frame_range[1], frame <= frame_range[2]) %>%
    transmute(frame, value = .data[[col1]], reliable = .data[[rel1]], instance = "inst1")
  
  df_all <- bind_rows(df0, df1)
  
  # plot
  ggplot(df_all, aes(x = frame, y = value, group = instance)) +
    geom_line(aes(color = instance), size = 0.5) +
    geom_point(data = df_all %>% filter(!reliable), aes(x = frame, y = value),
               color = "black", size = 0.8) +
    scale_color_manual(values = c("inst0" = "red", "inst1" = "blue")) +
    labs(title = paste(part, toupper(axis), "Frames", frame_range[1], "-", frame_range[2])) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
}

# --- Plot all parts and axes ---
parts <- c("mid", "nose", "tail")
axes  <- c("x", "y")

for (p in parts) {
  for (a in axes) {
    print(plot_part_axis(res, p, a))
  }
}

# --- per-part summary (unchanged) ---
reliability_summary <- function(res, part) {
  tibble(
    part = part,
    inst0_reliable   = sum(res[[paste0("reliability_", part, "0")]], na.rm = TRUE),
    inst0_unreliable = sum(!res[[paste0("reliability_", part, "0")]], na.rm = TRUE),
    inst1_reliable   = sum(res[[paste0("reliability_", part, "1")]], na.rm = TRUE),
    inst1_unreliable = sum(!res[[paste0("reliability_", part, "1")]], na.rm = TRUE)
  )
}

reliability_tables <- bind_rows(lapply(parts, reliability_summary, res = res))

# --- overall all/none reliable by instance ---
overall_reliability <- tibble(
  part = "ALL",
  inst0_reliable   = sum(res$reliability_mid0 & res$reliability_nose0 & res$reliability_tail0, na.rm = TRUE),
  inst0_unreliable = sum(!res$reliability_mid0 & !res$reliability_nose0 & !res$reliability_tail0, na.rm = TRUE),
  inst1_reliable   = sum(res$reliability_mid1 & res$reliability_nose1 & res$reliability_tail1, na.rm = TRUE),
  inst1_unreliable = sum(!res$reliability_mid1 & !res$reliability_nose1 & !res$reliability_tail1, na.rm = TRUE)
)

# combine into one table
reliability_tables <- bind_rows(reliability_tables, overall_reliability)

print(reliability_tables)


```

```{r}
# Create reliable-only data frames for each body part × instance
nose0_reliable_df <- res %>% filter(reliability_nose0)
mid0_reliable_df  <- res %>% filter(reliability_mid0)
tail0_reliable_df <- res %>% filter(reliability_tail0)

nose1_reliable_df <- res %>% filter(reliability_nose1)
mid1_reliable_df  <- res %>% filter(reliability_mid1)
tail1_reliable_df <- res %>% filter(reliability_tail1)

```

After the first phase, around 3/4 of the frames can be confirmed as reliable. Among the rest, some can be easily confirmed/extrapolated with nearby frames, while some have longer empty periods will need more advanced techniques to fill in such as Kalman filters and posterior density. Kalman filter both forwards and backwards can become a smoother(RTS) to predict the unreliable frames. To maintain conjugacy, the prior for the skeleton would need to be composed of the conditional Gaussian distribution based on the skeleton, as well as the Kalman filter from previous time points. The likelihood would then be based on the observed other body parts from the current frame.

Animation plot for frames 5000-6000 of the two mice, unreliable points colored as black:

```{r}
res_1000 <- res %>% filter(frame >= 4000, frame <= 5000)

# --- Pivot coordinates for instance 0 ---
coords0 <- res_1000 %>%
  dplyr::select(frame,
         inst0_nose_x, inst0_nose_y,
         inst0_mid_x,  inst0_mid_y,
         inst0_tail_x, inst0_tail_y) %>%
  pivot_longer(
    cols = -frame,
    names_to = c("body_part", ".value"),
    names_pattern = "inst0_(nose|mid|tail)_(x|y)"
  ) %>%
  mutate(body_part = dplyr::recode(body_part,
                            nose = "Nose",
                            mid  = "Mid-center",
                            tail = "Tail-base"),
         instance = "Instance 0")

# --- Pivot reliabilities for instance 0 ---
reliab0 <- res_1000 %>%
  dplyr::select(frame,
         reliability_nose0,
         reliability_mid0,
         reliability_tail0) %>%
  pivot_longer(
    cols = -frame,
    names_to = "body_part",
    names_pattern = "reliability_(nose|mid|tail)0"
  ) %>%
  rename(reliable = value) %>%
  mutate(body_part = dplyr::recode(body_part,
                            nose = "Nose",
                            mid  = "Mid-center",
                            tail = "Tail-base"),
         instance = "Instance 0")

# --- Pivot coordinates for instance 1 ---
coords1 <- res_1000 %>%
  dplyr::select(frame,
         inst1_nose_x, inst1_nose_y,
         inst1_mid_x,  inst1_mid_y,
         inst1_tail_x, inst1_tail_y) %>%
  pivot_longer(
    cols = -frame,
    names_to = c("body_part", ".value"),
    names_pattern = "inst1_(nose|mid|tail)_(x|y)"
  ) %>%
  mutate(body_part = dplyr::recode(body_part,
                            nose = "Nose",
                            mid  = "Mid-center",
                            tail = "Tail-base"),
         instance = "Instance 1")

# --- Pivot reliabilities for instance 1 ---
reliab1 <- res_1000 %>%
  dplyr::select(frame,
         reliability_nose1,
         reliability_mid1,
         reliability_tail1) %>%
  pivot_longer(
    cols = -frame,
    names_to = "body_part",
    names_pattern = "reliability_(nose|mid|tail)1"
  ) %>%
  rename(reliable = value) %>%
  mutate(body_part = dplyr::recode(body_part,
                            nose = "Nose",
                            mid  = "Mid-center",
                            tail = "Tail-base"),
         instance = "Instance 1")

# --- Combine instances ---
coords <- bind_rows(coords0, coords1)
reliab <- bind_rows(reliab0, reliab1)
df_anim <- left_join(coords, reliab,
                     by = c("frame", "body_part", "instance"))

# --- Plot ---
fig <- plot_ly(df_anim %>% filter(reliable, instance == "Instance 0"),
               x = ~x, y = ~y, frame = ~frame,
               color = ~body_part,
               colors = c("Nose" = "red",
                          "Mid-center" = "blue",
                          "Tail-base" = "green"),
               type = 'scatter', mode = 'markers+lines',
               line = list(width = 2),
               marker = list(symbol = "circle", size = 15)) %>%
  add_trace(data = df_anim %>% filter(!reliable, instance == "Instance 0"),
            x = ~x, y = ~y, frame = ~frame,
            type = 'scatter', mode = 'markers',
            marker = list(symbol = "circle", size = 15, color = "black"),
            inherit = FALSE,
            showlegend = FALSE) %>%
  add_trace(data = df_anim %>% filter(reliable, instance == "Instance 1"),
            x = ~x, y = ~y, frame = ~frame,
            color = ~body_part,
            colors = c("Nose" = "red",
                       "Mid-center" = "blue",
                       "Tail-base" = "green"),
            type = 'scatter', mode = 'markers+lines',
            line = list(width = 2),
            marker = list(symbol = "triangle-up", size = 15),
            inherit = FALSE) %>%
  add_trace(data = df_anim %>% filter(!reliable, instance == "Instance 1"),
            x = ~x, y = ~y, frame = ~frame,
            type = 'scatter', mode = 'markers',
            marker = list(symbol = "triangle-up", size = 15, color = "black"),
            inherit = FALSE,
            showlegend = FALSE) %>%
  layout(
    title = "Instances 0 (circle) and 1 (triangle) with reliability (black = unreliable)",
    xaxis = list(title = "X"),
    yaxis = list(title = "Y", scaleanchor = "x", autorange = "reversed")
  ) %>%
  animation_opts(
    frame = 40,  # ms per frame
    transition = 0,
    redraw = FALSE
  )

fig

```

```{r}
# For result: full counts of reliability
table(res$reliability_mid1)
table(res$reliability_mid0)
saveRDS(res, file = "Tracked.rds")
```

After the first phase, around 2/3 of the frames can be confirmed as reliable. Among the rest, some can be easily confirmed/extrapolated with nearby frames, while some have longer empty periods will need more advanced techniques to fill in such as Kalman filters and posterior density. Kalman filter both forwards and backwards can become a smoother(RTS) to predict the unreliable frames. To maintain conjugacy, the prior for the skeleton would need to be composed of the conditional Gaussian distribution based on the skeleton, as well as the Kalman filter from previous time points. The likelihood would then be based on the observed other body parts from the current frame.

Main problems:

First method gets high percentage of reliable frames for a single instance of a single body part, but it treats the two instances asymmetrically and causes wrong acceptances.

Second method uses negative log-likelihood score for each combination of instances for all body parts at once, but this is too strict and gets too low percentage of reliable body parts, along with many cases when a single body part is unreliable and affects the rest. This seems most promising so far, I can try to increase the accept rate but if there's no better choice, I will use this for the next step of Kalman filtering.

The third method uses Hungarian matching of instances to get a reliable track for a single anchor body part (e.g. mid-center), and it is generalizable for more instances and body parts. Currently, the biggest challenger is how to enforce the skeleton structure in this method, currently it can only handle body parts separately and causes bad consistency of skeletons.
